Compositional equivalences based on Open pNets

### 2021-10-05 2016-Weak-FH-Bisim/WeakBisim.tex
je pousse mes petites corrections de ce matin.

pNets is a generalisation of automata that model open systems. They features variables and hierarchical composition. Open pNets are pNets with holes, i.e. placeholders inside the hierarchical structure  that can be filled by sub-systems. 
%  Open pNets have a notion of \emph{synchronised actions}
%  generalizing the usual internal actions (e.g. $\tau$ of CCS, or $i$
%  in Lotos).

This article defines bisimulation relations  for the comparison of systems specified as pNets.
 We first define a strong bisimulation for open pNets.
  In practice, as happens in process algebras, strong bisimulation is too strong, and
  we define some coarser relations, taking into account
  invisible or internal moves.
  We then define an equivalence relation similar to
  the classical \emph{weak bisimulation}, and study its properties.
Among these properties we are interested in compositionality: If two systems are proven equivalent they will be undistinguishable by their context, and they will also be undistinguishable when their holes are filled with equivalent systems.
We identify sufficient conditions on the automata to ensure compositionality of strong and weak bisimulation.
The article is illustrated with a transport protocol running example; it shows the characteristics of our formalism and our bisimulation relations.
\end{abstract}

@@

There is also a clear relation with the seminal works on rule formats for Structured Operational Semantics, e.g. DeSimone format, GSOS, and conditional rules with or without negative premisses \cite{deSimone85,Bloom88bisimulationcant,GROOTE1992202,VANGLABBEEK2004229}.  Open pNets provides a way to define operators similar to these rules formats, but with quite different aim and approach. A formal comparison would be interesting, though not trivial. What we can say easily is that: the pNet format syntactically encompasses both DeSimone, GSOS, and conditional premisses rules. Then our compositionality result is more powerful than the classical result, but this is not a surprise, as it relies on a (sufficient) syntactic hypothesis on particular systems, rather than the general rules defining an operator.  Last, we intentionaly do not accept negative premisses, that would be more diffcult to put into practice in our implementation.

@@

\TODO{relire}
We are currently extending this work,  looking at further properties of FH-bisimulation, but also
the relations with existing equivalences on both closed and open systems. In particular, our model being significantly different from those considered in \cite{IngolfsdottirL:2001}, it
would be interesting to compare our "FH" family of bisimulations with the hierarchy of symbolic bisimulations from these authors.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have developed tool support for computing the
symbolic semantics in term of open automata \cite{QBMZ-AVOCS18}, and have
developed algorithms to check strong FH-bisimulation \cite{hou:hal-02406098}.
More reecntly we published preliminary work for the case of weak FH-Bisimulation \cite{}. 
The challenges here, in the context of our symbolic systems, is not so much algorithmic complexity,
as was the case with classical weak bisimulation on finite models, but decidability and termination.
The naive approach using an explicit construction of the weak transition, may in itself introduce non-termination,
so we prefer a direct implementation of the weak bisimulation definition, without
constructing the weak automaton, but searching \emph{on demand} to construct the required weak transitions.
Alternatively, we explore more pragmatic approaches using weak bisimulation preserving (pattern-based) 
reduction rules \cite{wang:hal-03126313}.

>>>>

pNets is a generalisation of automata that model open systems. They feature variables and hierarchical composition. Open pNets are pNets with holes, i.e. placeholders inside the hierarchical structure  that can be filled later by sub-systems. 
%  Open pNets have a notion of \emph{synchronised actions}
%  generalizing the usual internal actions (e.g. $\tau$ of CCS, or $i$
%  in Lotos).

This article defines bisimulation relations  for the comparison of systems specified as pNets.
 We first define a strong bisimulation for open pNets.
 \TODO{virer cette phrase, doublon du 1er paragraph?} In practice, as happens in process algebras, strong bisimulation is too strong, and
  we define some coarser relations, taking into account
  invisible or internal moves.
  We then define an equivalence relation similar to
  the classical \emph{weak bisimulation}, and study its properties.
Among these properties we are interested in compositionality: if two systems are proven equivalent they will be undistinguishable by their context, and they will also be undistinguishable when their holes are filled with equivalent systems.
We identify sufficient conditions on the automata to ensure compositionality of strong and weak bisimulation.
The article is illustrated with a transport protocol running example; it shows the characteristics of our formalism and our bisimulation relations.
\end{abstract}

@@

There is also a clear relation with the seminal works on rule formats for Structured Operational Semantics, e.g. DeSimone format, GSOS, and conditional rules with or without negative premisses \cite{deSimone85,Bloom88bisimulationcant,GROOTE1992202,VANGLABBEEK2004229}.  The Open pNets model provides a way to define operators similar to these rules formats, but with quite different aim and approach. A formal comparison would be interesting, though not trivial. What we can say easily is that: the pNet format syntactically encompasses both DeSimone, GSOS, and conditional premisses rules. Then our compositionality result is more powerful than their classical results, but this is not a surprise, as we rely on a (sufficient) syntactic hypothesis on a particular system, rather than the general rules defining an operator.  Last, we intentionaly do not accept negative premisses, that would be more diffcult to put into practice in our implementation. an extension could be studied in future work.

@@

\TODO{relire}
We are currently extending this work,  looking at further properties of FH-bisimulation, but also
the relations with existing equivalences on both closed and open systems. In particular, our model being significantly different from those considered in \cite{IngolfsdottirL:2001}, it
would be interesting to compare our "FH" family of bisimulations with the hierarchy of symbolic bisimulations from these authors.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have developed tool support for computing the
symbolic semantics in term of open automata \cite{QBMZ-AVOCS18}, and have
developed algorithms to check strong FH-bisimulation \cite{hou:hal-02406098}.
More recently we published preliminary work for the case of weak FH-Bisimulation \cite{}. 
The challenges here, in the context of our symbolic systems, is not so much algorithmic complexity,
as was the case with classical weak bisimulation on finite models, but decidability and termination.
The naive approach using an explicit construction of the weak transition, may in itself introduce non-termination,
so we prefer a direct implementation of the weak bisimulation definition, without
constructing the weak automata, but searching \emph{on demand} to construct the required weak transitions.
Beside, we explore in \cite{wang:hal-03126313} more pragmatic approaches using weak bisimulation preserving (pattern-based) 
reduction rules.
